Limited predictability of amino acid substitutions in seasonal influenza viruses

### 2020-12-16 article.tex
Downtone claims on p8 l47

Predicting the trajectory of a mutation requires (i) significant fitness difference between genomes carrying different variants at the site and (ii) a selection pressure that changes slowly over time.
Under such conditions, it is expected that frequency trajectories will show a persistent behavior which would make them predictable for some time.
However, we could find only limited evidence for such persistent behavior in the past 19 years of IAV evolution.
This lead us to conclude that (i) influenza virus evolution is qualitatively different from models of rapidly adapting population (despite clear evidence for frequent positive selection), and (ii) previous methods to predict influenza evolution work primarily because they pick strains that represent the future well, not because they predict future dynamics.

>>>>

Predicting the trajectory of a mutation requires (i) significant fitness difference between genomes carrying different variants at the site and (ii) a selection pressure that changes slowly over time.
Under such conditions, it is expected that frequency trajectories will show a persistent behavior which would make them predictable for some time.
However, we could find only limited evidence for such persistent behavior in the past 19 years of IAV evolution.
This lead us to conclude that (i) influenza virus evolution is qualitatively different from models of rapidly adapting population (despite clear evidence for frequent positive selection), and (ii) most methods to predict influenza evolution are unable to predict future frequency dynamics even though they work well for predicting the composition of the future population. 
This is illustrated by the case of LBI, which works primarily because it picks strains that represent the future well and not because it is predictive of future dynamics.
% work primarily because they pick strains that represent the future well, not because they predict future dynamics.
